---
title: "Power plant Descriptive Analysis"
author: "Roqeeb Olaniyi"
date: "14/06/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Objective of analysis
The power plant analysis was performed to answer the task questions below

* What is the mean value of daily yield?
* What is the total irradiation per day?
* What is the max ambient and module temperature?
* How many inverters are there for each plant?
* What is the maximum/minimum amount of DC/AC Power generated in a time interval/day?
* Which inverter (source_key) has produced maximum DC/AC power?
* Rank the inverters based on the DC/AC power they produce
* Is there any missing data?


Loading the libraries to be used for the analysis
```{r cars}
library(tidyverse)
library(ggplot2)
library(lubridate)
```
Reading the csv files to be used for the analysis
```{r}
plant_1_gen <- read_csv("C:/Users/ayoro/Desktop/Datasci/Case studies/Power plant/Plant_1_Generation_Data.csv")
plant_2_gen <- read_csv("C:/Users/ayoro/Desktop/Datasci/Case studies/Power plant/Plant_2_Generation_Data.csv")
plant_1_weather <- read_csv("C:/Users/ayoro/Desktop/Datasci/Case studies/Power plant/Plant_1_Weather_Sensor_Data.csv")
plant_2_weather <- read_csv("C:/Users/ayoro/Desktop/Datasci/Case studies/Power plant/Plant_2_Weather_Sensor_Data.csv")
```

Initial data exploration
```{r}
head(plant_1_gen)
```

```{r}
glimpse(plant_1_gen)
```

```{r}
glimpse(plant_2_gen)
```

```{r}
head(plant_1_weather)
```

```{r}
glimpse(plant_1_weather)
```

```{r}
glimpse(plant_2_weather)
```

Checking for null or empty values in column of each dataframe
```{r}
colSums(is.na(plant_1_gen))
colSums(is.na(plant_2_gen))
colSums(is.na(plant_1_weather))
colSums(is.na(plant_2_weather))
```

Creating a date column from the date_time column to make analysis better
```{r}
plant_1_gen <- plant_1_gen %>%
                    mutate('date' = date(dmy_hm(DATE_TIME)))

plant_2_gen <- plant_2_gen %>%
                    mutate('date' = date(DATE_TIME))

plant_1_weather <- plant_1_weather %>%
                    mutate('date' = date(DATE_TIME))

plant_2_weather <- plant_2_weather %>%
                    mutate('date' = date(DATE_TIME))
```

Getting mean daily yield for each plant
```{r}
plant_1_gen %>%
summarise(mean = mean(DAILY_YIELD))
```

```{r}
plant_2_gen %>%
summarise(mean = mean(DAILY_YIELD))
```

#### Total irradiation per day 

Plant 1
```{r}
plant_1_weather %>%
group_by(date) %>%
summarise('Total_irradiation' = sum(IRRADIATION)) %>%
  
  ggplot(aes(x = date, y = Total_irradiation)) +
  geom_line() +
  labs(title = 'Irradiation per day', y = 'Total Irradiation', x = 'Date') +
  theme_minimal() +
  theme(panel.grid = element_blank())
```

Plant 2
```{r}
plant_2_weather %>%
group_by(date) %>%
summarise('Total_irradiation' = sum(IRRADIATION)) %>%
  ggplot(aes(x = date, y = Total_irradiation)) +
  geom_line() +
  labs(title = 'Irradiation per day', y = 'Total Irradiation', x = 'Date') +
  theme_minimal() +
  theme(panel.grid = element_blank())
```

#### Max ambient and module temperature

Plant 1 
```{r}
plant_1_weather %>%
summarise('max_ambient_temp' = max(AMBIENT_TEMPERATURE), 'max_module_temp' = max(MODULE_TEMPERATURE))
```

Plant 2
```{r}
plant_2_weather %>%
summarise('max_ambient_temp' = max(AMBIENT_TEMPERATURE), 'max_module_temp' = max(MODULE_TEMPERATURE))
```

#### Number of inverters

Plant 1
```{r}
plant_1_gen %>%
summarize('inverters' = n_distinct(SOURCE_KEY))
```

Plant 2
```{r}
plant_2_gen %>%
summarize('inverters' = n_distinct(SOURCE_KEY))
```

#### Maximum and minumum AC/DC power per day

Plant 1
```{r}
max_AC_DC_1 <- plant_1_gen %>%
                group_by(date) %>%
                summarise('max_AC' = max(AC_POWER),
                'min_AC' = min(AC_POWER),
                'max_DC' = max(DC_POWER),
                'min_DC' = min(DC_POWER))

max_AC_DC_1 %>%
  ggplot(aes(x = date, y = max_AC)) +
  geom_line() +
  labs(title = 'Maximum AC power per day', x = 'Date', y = 'AC (kw)') +
  theme_minimal() +
  theme(panel.grid = element_blank())

max_AC_DC_1 %>%
  ggplot(aes(x = date, y = max_DC)) +
  geom_line()  +
  labs(title = 'Maximum DC power per day', x = 'Date', y = 'DC (kw)') +
  theme_minimal() +
  theme(panel.grid = element_blank())
```

Plant 2
```{r}
max_AC_DC_2 <- plant_2_gen %>%
              group_by(date) %>%
              summarise('max_AC' = max(AC_POWER),
              'min_AC' = min(AC_POWER),
              'max_DC' = max(DC_POWER),
              'min_DC' = min(DC_POWER))
max_AC_DC_2 %>%
  ggplot(aes(x = date, y = max_AC)) +
  geom_line() +
  labs(title = 'Maximum AC power per day', x = 'Date', y = 'AC (kw)') +
  theme_minimal() +
  theme(panel.grid = element_blank())

max_AC_DC_2 %>%
  ggplot(aes(x = date, y = max_AC)) +
  geom_line() +
  labs(title = 'Maximum DC power per day', x = 'Date', y = 'DC (kw)') +
  theme_minimal() +
  theme(panel.grid = element_blank())
```


#### Inverter with the maximum AC/DC power output

Plant 1
```{r}
plant_1_gen %>%
filter(AC_POWER == max(AC_POWER) | DC_POWER == max(DC_POWER)) %>%
select('max AC/DC inverter' = SOURCE_KEY)
```

Plant 2
```{r}
plant_2_gen %>%
filter(AC_POWER == max(AC_POWER) | DC_POWER == max(DC_POWER)) %>%
select('max AC/DC inverter' = SOURCE_KEY)
```


#### Inverter rank based on AC/DC power output

Plant 1
```{r}
plant_1_gen %>%
group_by(SOURCE_KEY) %>%
select(SOURCE_KEY, AC_POWER, DC_POWER) %>%
summarise('AC_POWER' = sum(AC_POWER), 'DC_POWER' = sum(DC_POWER)) %>%
arrange(desc(AC_POWER, DC_POWER)) %>%
mutate('rank' = dense_rank(interaction(desc(AC_POWER),desc(DC_POWER)))) %>%
select(rank,SOURCE_KEY, AC_POWER, DC_POWER)
```

Plant 2
```{r}
plant_2_gen %>%
group_by(SOURCE_KEY) %>%
select(SOURCE_KEY, AC_POWER, DC_POWER) %>%
summarise('AC_POWER' = sum(AC_POWER), 'DC_POWER' = sum(DC_POWER)) %>%
arrange(desc(AC_POWER, DC_POWER)) %>%
mutate('rank' = dense_rank(interaction(desc(AC_POWER),desc(DC_POWER)))) %>%
select(rank,SOURCE_KEY, AC_POWER, DC_POWER)
```











